---
title: "SAGE promoter analysis"
output: html_document
---
###############################################################################
 Code written by Joshua Elmore (joshua.elmore@pnnl.gov) for SAGE: Serine      #
 recombinase-assisted genome engineering manuscript                           #
 Note, this  worksheet is for processing non-log scale values.                #
###############################################################################

Below are the input variables for the function used to generate plots for
individual organisms and a summary of promoter stats for each organism.

##Const_promoter_stats_indiv_org <- function( Input_file_label,
##                                            Output_file_label,
##                                            num_cond,
##                                            num_bc_rep,
##                                            num_exp_rep,
##                                            UTR_cutoff
##                                            Cond_cutoff
##                                          )

Input_file_label is the label infront of the generic R input file name:
e.g. CGA009 in 
"CGA009_promoter_full_dataframe_NGS_cutoff5_3MAD_min3reps_for_R.csv"

Output_file_label defines the folder name used and file name generated
for printed charts (e.g. Rpal for R. palustris CGA009 data)

num_cond is the number of conditions tested in the experiment
defaults to 3 if not defined below

num_bc_reps is the number of barcodes per promoter used in the experiment
defaults to 5 if not defined below

num_exp_reps is the number of experimental replicates performed (e.g. 
number of cultures harvested and processed for each strain/condition)
defaults to 4 if not defined below

UTR_cut is the value used for the maximum fold-difference between barcodes 
within a promoter that will be used to classify a promoter UTR sensitive.
This defaults to 3 if not defined below

Cond_cut is the value used for the maximum fold-difference between conditions 
within a promoter that will be used to classify a promoter condition 
sensitive.
This defaults to 4 if not defined below

## Code chunk that loads the source files for each function.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("Const_prom_stats_ind_org.R")
source("Const_prom_stats_ind_org_median.R")
source("Const_prom_stats_ind_org_no_charts.R")
source("Const_prom_stats_org_comp_data.R")
source("generate_org_compare_charts-set1.R")
source("summarySE_median.R")

## Use these to define values used for UTR_cut and Cond_cut. If unused, the
## functions below will use 3 and 4 for these variables respectively.
UTR_cutoff = 3.5
Cond_cutoff = 4

```

## Code chunk that generates plots for each organism using mean promoter 
## strength for calculation, and generates a csv that contains several 
## general promoter stats for each organism

```{r individual_organism_plots, include=FALSE}
## Initialize an empty dataframe for the promoter analysis overview
promoter_analysis_overview <- data.frame(matrix(ncol = 13, nrow = 0))
## add column names to the dataframe
column_names <- c("Organism",
                  "num_bc_pass",
                  "num_prom_in_bc_pass",
                  "num_prom_suff_data",
                  "num_bc_pass_outlier",
                  "num_UTR_insensitive",
                  "num_Cond_insensitive",
                  "num_consistent",
                  "UTR_cutoff",
                  "Cond_cutoff",
                  "expression_range",
                  "Problematic_percent",
                  "Number_of_promoters_w_prob_bc"
                  )

## Generate a dataframe with a broad overview of stats from the promoter
## analysis. This dataframe is for the stats when using the default UTR and
## Condition insensitivity cutoffs of 3.5 and 4 respectively.
promoter_analysis_overview <- rbind(promoter_analysis_overview,
                                    Const_promoter_stats_indiv_org("RHA1",
                                                                   "Rjostii",
                                                                   3,
                                                                   5,
                                                                   4,
                                                                   UTR_cutoff,
                                                                   Cond_cutoff)
                                    )
promoter_analysis_overview <- rbind(promoter_analysis_overview, 
                                    Const_promoter_stats_indiv_org("Pflu",
                                                                   "Pflu",
                                                                   4,
                                                                   5,
                                                                   4,
                                                                   UTR_cutoff,
                                                                   Cond_cutoff)
                                    )
promoter_analysis_overview <- rbind(promoter_analysis_overview, 
                                    Const_promoter_stats_indiv_org("TBS10",
                                                                   "Pbr",
                                                                   3,
                                                                   5,
                                                                   4,
                                                                   UTR_cutoff,
                                                                   Cond_cutoff)
                                    )
promoter_analysis_overview <- rbind(promoter_analysis_overview, 
                                    Const_promoter_stats_indiv_org("Gpo1",
                                                                   "Ppu",
                                                                   3,
                                                                   5,
                                                                   4,
                                                                   UTR_cutoff,
                                                                   Cond_cutoff)
                                    )
promoter_analysis_overview <- rbind(promoter_analysis_overview, 
                                    Const_promoter_stats_indiv_org("CGA009",
                                                                   "Rpal",
                                                                   2,
                                                                   5,
                                                                   4,
                                                                   UTR_cutoff,
                                                                   Cond_cutoff)
                                    )
write_csv(promoter_analysis_overview,
          "./Promoter_stats/promoter_analysis_overview.csv")

```

## Code chunk that generates plots for each organism using median promoter 
## strength for calculation, and generates a csv that contains several 
## general promoter stats for each organism

```{r individual_organism_plots, include=FALSE}
## Initialize an empty dataframe for the promoter analysis overview
promoter_analysis_overview <- data.frame(matrix(ncol = 13, nrow = 0))
## add column names to the dataframe
column_names <- c("Organism",
                  "num_bc_pass",
                  "num_prom_in_bc_pass",
                  "num_prom_suff_data",
                  "num_bc_pass_outlier",
                  "num_UTR_insensitive",
                  "num_Cond_insensitive",
                  "num_consistent",
                  "UTR_cutoff",
                  "Cond_cutoff",
                  "expression_range",
                  "Problematic_percent"
                  )

## Generate a dataframe with a broad overview of stats from the promoter
## analysis. This dataframe is for the stats when using the default UTR and
## Condition insensitivity cutoffs of 3.5 and 4 respectively.
promoter_analysis_overview <- 
  rbind(promoter_analysis_overview,
        Const_promoter_stats_indiv_org_median("RHA1",
                                       "Rjostii",
                                       3,
                                       5,
                                       4,
                                       UTR_cutoff,
                                       Cond_cutoff)
                                    )
promoter_analysis_overview <- 
  rbind(promoter_analysis_overview, 
        Const_promoter_stats_indiv_org_median("Pflu",
                                       "Pflu",
                                       4,
                                       5,
                                       4,
                                       UTR_cutoff,
                                       Cond_cutoff)
                                    )
promoter_analysis_overview <- 
  rbind(promoter_analysis_overview, 
        Const_promoter_stats_indiv_org_median("TBS10",
                                       "Pbr",
                                       3,
                                       5,
                                       4,
                                       UTR_cutoff,
                                       Cond_cutoff)
                                    )
promoter_analysis_overview <- 
  rbind(promoter_analysis_overview, 
        Const_promoter_stats_indiv_org_median("Gpo1",
                                       "Ppu",
                                       3,
                                       5,
                                       4,
                                       UTR_cutoff,
                                       Cond_cutoff)
                                    )
promoter_analysis_overview <- 
  rbind(promoter_analysis_overview, 
        Const_promoter_stats_indiv_org_median("CGA009",
                                       "Rpal",
                                       2,
                                       5,
                                       4,
                                       UTR_cutoff,
                                       Cond_cutoff)
                                    )
write_csv(promoter_analysis_overview,
          "./Promoter_stats/promoter_analysis_overview_median.csv")

```

## Code chunk that generates csv's that contain all of the same promoter stats
## as generated above, but they do so for each organism using an array of
## inputed cutoffs, rather than just a single one as above.

```{r parameter_sweep, include = FALSE}
################################################################################
## Sweep of possible stats if different sensitivity cutoffs were used.        ##
## Runs nearly identical function as above, but this function does not        ##
## generate and write chart files.                                            ##
## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ##

## Running parameter sweep for Rhodococcus jostii RHA1
UTR_parameters <- c(2,2.5,3,3.5,4,5,6,7,8,10)
Cond_parameters <- c(2,2.5,3,3.5,4,5,6,7,8,10) 

parameter_sweep_RHA1 <- data.frame(matrix(ncol = 13, nrow = 0))
colnames(parameter_sweep_RHA1) <- column_names

  for(valU in UTR_parameters){
    for(valC in Cond_parameters){
      parameter_sweep_RHA1 <-rbind(parameter_sweep_RHA1, 
                              Const_prom_stats_ind_org_no_charts(
                                "RHA1",
                                "Rjostii",
                                3,
                                5,
                                4,
                                valU,
                                valC)
                              )
      
    }
  }
write_csv(parameter_sweep_RHA1,
          "./Promoter_stats/parameter sweep RHA1.csv")

## Running parameter sweep for Pseudomonas fluorescens SBW25
parameter_sweep_SBW25 <- data.frame(matrix(ncol = 13, nrow = 0))
colnames(parameter_sweep_SBW25) <- column_names

  for(valU in UTR_parameters){
    for(valC in Cond_parameters){
      parameter_sweep_SBW25 <-rbind(parameter_sweep_SBW25, 
                              Const_prom_stats_ind_org_no_charts(
                                "Pflu",
                                "Pflu",
                                4,
                                5,
                                4,
                                valU,
                                valC)
      )
      
    }
  }
write_csv(parameter_sweep_SBW25,
          "./Promoter_stats/parameter sweep SBW25.csv")

## Running parameter sweep for Pseudomonas putida Gpo1
parameter_sweep_Gpo1 <- data.frame(matrix(ncol = 13, nrow = 0))
colnames(parameter_sweep_Gpo1) <- column_names

for(valU in UTR_parameters){
  for(valC in Cond_parameters){
    parameter_sweep_Gpo1 <-rbind(parameter_sweep_Gpo1, 
                            Const_prom_stats_ind_org_no_charts(
                              "Gpo1",
                              "Ppu",
                              3,
                              5,
                              4,
                              valU,
                              valC)
    )
    
  }
}
write_csv(parameter_sweep_Gpo1,
          "./Promoter_stats/parameter sweep Gpo1.csv")

## Running parameter sweep for Pseudomonas brassicacaerum TBS10
parameter_sweep_TBS10 <- data.frame(matrix(ncol = 13, nrow = 0))
colnames(parameter_sweep_TBS10) <- column_names

for(valU in UTR_parameters){
  for(valC in Cond_parameters){
    parameter_sweep_TBS10 <-rbind(parameter_sweep_TBS10, 
                            Const_prom_stats_ind_org_no_charts(
                              "TBS10",
                              "Pbr",
                              3,
                              5,
                              4,
                              valU,
                              valC)
    )
    
  }
}
write_csv(parameter_sweep_TBS10,
          "./Promoter_stats/parameter sweep TBS10.csv")


## Running parameter sweep for Rhodopseudomonas palustris CGA009
parameter_sweep_CGA009 <- data.frame(matrix(ncol = 13, nrow = 0))
colnames(parameter_sweep_CGA009) <- column_names

for(valU in UTR_parameters){
  for(valC in Cond_parameters){
    parameter_sweep_CGA009 <-rbind(parameter_sweep_CGA009, 
                            Const_prom_stats_ind_org_no_charts(
                              "CGA009",
                              "Rpal",
                              2,
                              5,
                              4,
                              valU,
                              valC)
    )
    
  }
}
write_csv(parameter_sweep_CGA009,
          "./Promoter_stats/parameter sweep CGA009.csv")

```

## Code chunk that generates the dataframes needed to perform comparisons
## between strains.

## Below are the names and descriptions of the variables in the 
## all_stats_organisms.csv generated by the below code chunk
## each line in the document is for a specific barcode replicate in a given 
## physical sample (e.g. shake for replicate for different conditions, organism,
## etc.)
                Name =                 Name of promoter
                Replicate =            Barcode replicate number (values 1-5)
                UTR_sensitive =        Does the promoter have a 5'UTR senstivity
                UTR_max_min_ratio =    What is the ratio between the strongest 
                                       and weakest barcode variant for the 
                                       promoter
                Cond_sensitive =       Does the promoter display condition 
                                       sensitivity
                Cond_max_min_ratio =   What is the ratio between the conditions 
                                       displaying the strongest and weakest 
                                       expression level
                Overall_PS =           Mean RTA when all filtered 
                                       promoter strength values are averaged. 
                                       This filters out data points for barcodes 
                                       that either lack sufficient data for 
                                       analysis or are outliers. Unlike 
                                       Overall_PS_BC, no averaging is performed 
                                       as an intermediate step.
                Barcode =              barcode sequence for data point
                bc_outlier =           is this barcode an outlier? If so, this 
                                       data point was not used to calculate the 
                                       Overall_PS or Overall_PS_BC values 
                N_promoter_samples =   Number of data points for a given 
                                       promoter in the indcated organism
                N_barcode_samples =    Number of samples for the given barcode
                                       in the indicated organism
                Name_and_bc_rep =      Naming scheme that includes the promoter 
                                       name and barcode replicate # (values 1-5)
                Condition =            Condition tested
                Condition_replicate=   Replicate shake flask # (values 1-4)
                Promoter_Strength =    RTA value for the given sample.
                Sequence =             Promoter / barcode sequence
                consistent =           Label for is the promoter activity is 
                                       considered consistent in the given 
                                       organism. For this, the 
                                       Cond_min_max_ratio must be lower than the
                                       defined threshold (default 4) and 
                                       UTR_min_max_ratio lower than its defined 
                                       threshold (default 3.5) in the given 
                                       organism
                organism =             organism from which the data point was 
                                       derived
                BC_Promoter_Strength = mean RTA across conditions for the 
                                       indicated promoter barcode variant in the
                                       given organism
                Overall_PS_BC =        Same as Overall_PS, except it uses the 
                                       averages of BC_promoter_strength values 
                                       for non-outlier barcode variants to 
                                       determine the Overall RTA value. This 
                                       eliminates some barcode variants from 
                                       having more weight in the mean/medians 
                                       than others. Basically, if one barcode 
                                       has 12 data points and another has 16, 
                                       the latter will have a stronger influence 
                                       in the Overall_PS value, but not for this 
                                       value.

```{r generate_data_needed_for_strain_comparison, include = FALSE}
## Initialize an empty dataframe for the promoter comparison between organisms
organism_comparison <- data.frame(matrix(ncol = 20, nrow = 0))
column_names <- c("Name",
                  "Replicate",
                  "UTR_sensitive",
                  "UTR_max_min_ratio",
                  "Cond_sensitive",
                  "Cond_max_min_ratio",
                  "Overall_PS",
                  "Barcode",
                  "bc_outlier",
                  "N_promoter_samples",
                  "N_barcode_samples",
                  "Name_and_bc_rep",
                  "Condition",
                  "Condition_replicate",
                  "Promoter_Strength",
                  "Sequence",
                  "consistent",
                  "organism",
                  "BC_Promoter_Strength",
                  "Overall_PS_BC"
                  )
colnames(organism_comparison) <- column_names

## Note, select the organism that has the most barcodes represented in its 
## non-barcode outlier data to be the first here. It will come in handy later.
## This data can be found in the output of an above chunk of code (the file:
## "promoter_analysis_overview.csv"). There is likely a way to populate these 
##  in order by factoring the data in that table by the variable in question, 
## and then using it to generate a for loop that runs eaech of the following 
## functions. However, this code will rely on manual rearrangement.

organism_comparison <- rbind(organism_comparison, 
                                    Const_prom_stats_org_comp_data("TBS10",
                                                                   "Pbr",
                                                                   3,
                                                                   5,
                                                                   4)
                                    )
organism_comparison <- rbind(organism_comparison,
                                   Const_prom_stats_org_comp_data("RHA1",
                                                                   "Rjostii",
                                                                   3,
                                                                   5,
                                                                   4)
                                    )
organism_comparison <- rbind(organism_comparison, 
                                    Const_prom_stats_org_comp_data("Pflu",
                                                                   "Pflu",
                                                                   4,
                                                                   5,
                                                                   4)
                                    )

organism_comparison <- rbind(organism_comparison, 
                                    Const_prom_stats_org_comp_data("Gpo1",
                                                                   "Ppu",
                                                                   3,
                                                                   5,
                                                                   4)
                                    )
organism_comparison <- rbind(organism_comparison, 
                                    Const_prom_stats_org_comp_data("CGA009",
                                                                   "Rpal",
                                                                   2,
                                                                   5,
                                                                   4)
                                    )

## generate dataframe with promoter strength data that is seperated by both 
## barcode and organism. Barcodes that were outliers are either included 
## (PS_org_and_bc_w_out) or excluded (PS_org_and_bc)
PS_org_and_bc_w_out <- organism_comparison

PS_org_and_bc_w_out <- subset(PS_org_and_bc_w_out, select = c("Name",
                                                        "organism",
                                                        "Overall_PS",
                                                        "Replicate",
                                                        "BC_Promoter_Strength",
                                                        "bc_outlier",
                                                        "consistent",
                                                        "Overall_PS_BC"))
PS_org_and_bc <- PS_org_and_bc_w_out %>% 
  filter(PS_org_and_bc_w_out$bc_outlier == "No" )

## remove duplicated rows
PS_org_and_bc_w_out <- unique(PS_org_and_bc_w_out)
PS_org_and_bc <- unique(PS_org_and_bc)
PS_org_and_bc_consistent <-PS_org_and_bc %>%
  filter(PS_org_and_bc$consistent == "(1) yes")

## Generate an array containing the distinct organisms used in the experiment.
## This will be used to filter a data frame into subsets of the data containing
## only the data for that condition
  organism_array <- unique(subset(PS_org_and_bc_w_out, 
                                   select = c("organism")
                                   )
                            )

## generate dataframe with promoter strength data that is seperated just by 
## organism (all barcodes compressed)
  
write_csv(organism_comparison,
          "./Promoter_stats/all_organism_stats.csv")

```

```{r all_by_all_organism_comparisons_chart1, include = FALSE}


## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ##
## Generate a scatter plot comparing performance of individual promoters across
## each organism. Dots are mean Promoter_Strength for a given organism, and are
## colored according to organism. A regression line will be generated for each
## organism to highlight both the general trend in promoter strengths across
## organism. The steepness of the regression line slope will also serve as an
## indicator of the dynamic range of promoter performance in a given organism.
## Generally, a shallow slope (sub 1) means there is less difference in promoter
## strength from strongest to weakest promoter for the x-axis condition than for
## the y-axis condition. The converse is true for a >1 slope. If all slopes are
## close to 1, it will suggest that the relative performance of the promoters
## in the library are fairly similar between organisms. 
## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ## ##


## Generate a subset of the PS_org_and_bc that pulls out the first
## organism in the array. Doing this allows us to set it as the X-axis in 
## promoter comparison charts

  d_org_1 <- subset(PS_org_and_bc_consistent, 
                     organism == organism_array$organism[1], 
                     select = c("Name",
                                "organism",
                                "Overall_PS",
                                "Overall_PS_BC"))
    
## These generate a distinct label for the promoter strength and standard error
## for the organism excised in the previous line of code. Together these 
## lines allow us to keep the data for this organism as a distinct variable
## when we recombine all of the organism data into a single reformatted
## dataframe below
  org_overall_PS_lbl = paste(organism_array$organism[1], 
                             "_Overall_PS", 
                             sep="")
  
  org_BC_PS_label = paste(organism_array$organism[1],  
                          "_BC_PS", 
                          sep="")
  
  org_overall_PS_BC_lbl = paste(organism_array$organism[1], 
                                "_Overall_PS_BC", 
                                sep="")
  
  
  colnames(d_org_1)<-c("Name",
                       "organism", 
                       org_overall_PS_lbl, 
                       org_overall_PS_BC_lbl)

## Generate a subset of the PS_org_and_bc that pulls out the all condition
## condition in the array that is not in the first condition in the array. 
  d_org_others <- subset(PS_org_and_bc_consistent, 
                         organism != organism_array$organism[1], 
                         select = c("Name",
                                    "organism",
                                    "Overall_PS",
                                    "Overall_PS_BC"))

## These generate a distinct label for the promoter strength and standard error
## for the organisms excised in the previous line of code. Together these three
## lines allow us to keep the data for this organisms as a distinct variable
## when we recombine all of the organism data into a single reformatted
## dataframe below  
  org_overall_PS_lbl2 = paste("not_",
                            organism_array$organism[1], 
                            "_Overall_PS", 
                            sep="")
  
  org_BC_PS_label2 = paste("not_",
                          organism_array$organism[1],  
                          "_BC_PS", 
                          sep="")
  
  org_overall_PS_BC_lbl2 = paste("not",
                                organism_array$organism[1], 
                                "_Overall_PS_BC", 
                                 sep="")
  
  colnames(d_org_others)<-c("Name",
                            "organism", 
                            org_overall_PS_lbl2, 
                            org_overall_PS_BC_lbl2)


## merge the dataframes generated above. Merging these two allows us to use the
## single dataframe for charting promoter strengths for one organism on x-axis
## and promoter strengths for the other orgqnisms on the y-axis with different
## colored markers for each organism
  d_org_comp <- merge(d_org_1, d_org_others, by = "Name")


print_me <-  generate_org_compare_charts_1 (d_org_comp,
                                          organism_array$organism[1],
                                          "all others")


```

```{r pair_wise_organism_comparison_charts1, include = FALSE}

## Generate charts as in the previous chunk, but instead of a single chart with
## all organisms, generate a collection of all pair-wise comparisons between 
## organisms

valX <- 1
valY <- 1
while(valX <= nrow(organism_array)){
  while (valY <= nrow(organism_array)) {
    if(valX != valY){
    ## Generate a subset of the PS_org_and_bc that pulls out the organisms valX
    ## and valY in the array. Doing this allows us to set them as the X-axis and
    ## Y-axis in promoter comparison charts
    d_orgX <- subset(PS_org_and_bc, 
                     organism == organism_array$organism[valX], 
                     select = c("Name",
                                "organism",
                                "Overall_PS",
                                "Overall_PS_BC"))
    
    d_orgY <- subset(PS_org_and_bc, 
                     organism == organism_array$organism[valY], 
                     select = c("Name",
                                "organism",
                                "Overall_PS",
                                "Overall_PS_BC"))
    ## These generate a distinct label for the promoter strength and standard
    ## error for the organism excised in the previous line of code. Together
    ## these lines allow us to keep the data for this organism as a distinct
    ## variable when we recombine all of the organism data into a single
    ## reformatted dataframe below
    X_PS_lbl = paste(organism_array$organism[valX], "_Overall_PS", sep="")
    X_PS_BC_lbl = paste(organism_array$organism[valY], "_Overall_PS_BC", sep="")
    colnames(d_orgX)<-c("Name", "organism", X_PS_lbl, X_PS_BC_lbl)
    
    Y_PS_lbl = paste(organism_array$organism[valY], "_Overall_PS", sep="")
    Y_PS_BC_lbl = paste(organism_array$organism[valY], "_Overall_PS_BC", sep="")
    colnames(d_orgY)<-c("Name", "organism", Y_PS_lbl, Y_PS_BC_lbl)

    ## merge the dataframes generated above. Merging these two allows us to use
    ## the single dataframe for charting promoter strengths for one organism on
    ## x-axis and promoter strengths for the other orgqnisms on the y-axis with
    ## different colored markers for each organism
    d_org_XY <- merge(d_orgX, d_orgY, by = "Name")
    print_me <- generate_org_compare_charts_1 (d_org_XY,
                                             organism_array$organism[valX],
                                             organism_array$organism[valY],
                                             )
    plot(print_me, width = 5, height = 4)
    }
    valY <- (valY + 1)
  } 
  valY <- 1
  valX <- (valX + 1)
}

```

```{r consistent_pair_wise_organism_comparison_charts1, include = FALSE}

## Generate charts as in the previous chunk, but instead of a single chart with
## all organisms, generate a collection of all pair-wise comparisons between 
## organisms

valX <- 1
valY <- 1
while(valX <= nrow(organism_array)){
  while (valY <= nrow(organism_array)) {
    if(valX != valY){
    ## Generate a subset of the PS_org_and_bc that pulls out the organisms valX
    ## and valY in the array. Doing this allows us to set them as the X-axis and
    ## Y-axis in promoter comparison charts
    d_orgX <- subset(PS_org_and_bc_consistent, 
                     organism == organism_array$organism[valX], 
                     select = c("Name",
                                "organism",
                                "Overall_PS",
                                "Overall_PS_BC"))
    
    d_orgY <- subset(PS_org_and_bc_consistent, 
                     organism == organism_array$organism[valY], 
                     select = c("Name",
                                "organism",
                                "Overall_PS",
                                "Overall_PS_BC"))
    ## These generate a distinct label for the promoter strength and standard
    ## error for the organism excised in the previous line of code. Together
    ## these lines allow us to keep the data for this organism as a distinct
    ## variable when we recombine all of the organism data into a single
    ## reformatted dataframe below
    X_PS_lbl = paste(organism_array$organism[valX], "_Overall_PS", sep="")
    X_PS_BC_lbl = paste(organism_array$organism[valY], "_Overall_PS_BC", sep="")
    colnames(d_orgX)<-c("Name", "organism", X_PS_lbl, X_PS_BC_lbl)
    
    Y_PS_lbl = paste(organism_array$organism[valY], "_Overall_PS", sep="")
    Y_PS_BC_lbl = paste(organism_array$organism[valY], "_Overall_PS_BC", sep="")
    colnames(d_orgY)<-c("Name", "organism", Y_PS_lbl, Y_PS_BC_lbl)

    ## merge the dataframes generated above. Merging these two allows us to use
    ## the single dataframe for charting promoter strengths for one organism on
    ## x-axis and promoter strengths for the other orgqnisms on the y-axis with
    ## different colored markers for each organism
    d_org_XY <- merge(d_orgX, d_orgY, by = "Name")
    print_me <- generate_org_compare_charts_1 (d_org_XY,
                                             organism_array$organism[valX],
                                             organism_array$organism[valY],
                                             )
    plot(print_me, width = 5, height = 4)
    }
    valY <- (valY + 1)
  } 
  valY <- 1
  valX <- (valX + 1)
}

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent 
printing of the R code that generated the plot.
